---
title: "Vignette MaMuT2R"
author: "Marion Louveaux"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE}
library(MaMuT2R)
```


## Load .xml MaMuT file

```{r, echo=FALSE}
MaMuTpath <- "/Users/mlouveaux/Documents/Postdoc/Softwares/Fiji/Test_plugins/2017-04-26_MaMuT/MaMuT_Parhyale_demo-mamut.xml"

MaMuT_XML <- XML::xmlToList(MaMuTpath)
```


## Get information relative to either nuclei (= spots) or lineages (=tracks)

MaMuT .xml file stores two type of objects, under two types of tags: the spots and the tracks. In the case of a biological sample, spots usually refer to nuclei and tracks to lineages. Here, Spots.as.dataframe() and Tracks.as.dataframe() extract information from .xml file and format them as dataframes.

```{r, echo=FALSE}
Spots_df <- Spots.as.dataframe(MaMuT_XML)
Tracks_df <- Tracks.as.dataframe(MaMuT_XML)
```







