---
title: "Vignette MaMuT2R"
author: "Marion Louveaux"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo=FALSE}
library(MaMuT2R)
```


## Load .xml MaMuT file

```{r, echo=FALSE}
XML_dir <- "/Users/mlouveaux/Documents/Postdoc/Projects/Nuclei_segmentation/2017-05-29_MaMuT_H5"
XML_file <- "2017-02-23_16.59.21_stPVB003-2-2xDR5v2_test_cropped_dc_export-mamut.xml"
MaMuTpath <- paste0(XML_dir, "/", XML_file)
MaMuT_XML <- readMaMuT(MaMuTpath)
```


## Get information relative to either nuclei (= spots) or lineages (=tracks)

MaMuT .xml file stores two type of objects, under two types of tags: the spots and the tracks. In the case of a biological sample, spots usually refer to nuclei and tracks to lineages. Here, Spots.as.dataframe() and Tracks.as.dataframe() extract information from .xml file and format them as dataframes.

```{r, echo=FALSE}
Spots_df <- Spots.as.dataframe(MaMuT_XML)
Tracks_df <- Tracks.as.dataframe(MaMuT_XML)
```



